---
title: "descriptiveAnalysis"
author: "Benjamin Carter"
date: "9/23/2020"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
# ENVIRONMENT ####
# packages
library(dplyr)
library(tidyr)
library(lubridate)
library(icd) # for reading in ICD10 classifying information and calculating risk scores
library(ggplot2)
library(compareGroups)

```

## Primary hypothesis

If Same day care readmissions are predictable then they can be prevented or curtailed.

## Aims

The aims of this analysis are as follows
- How much traffic are we seeing in SDC/EC?
- What are the top ICD-10 codes associated with a readmission?
- Who is coming back?
- How many times are they coming back?
- How often are they coming back?
- Why are they coming back?
- Where/who are they returning too?

## SDC Traffic

```{r, echo=FALSE, warning=FALSE}
fliers <- df.flat %>% 
  group_by(
    PersonID
  ) %>% 
  summarise(
    n = n()
  ) %>% 
  ungroup()

fliers %>% 
  ggplot(
    aes(n)
  ) +
  geom_histogram() +
  scale_y_log10() +
  theme_classic() +
  labs(
    title = "Distribution of Average Number of Visits to SDC/EC",
    x = "Number of Visits",
    y = "Number of Individuals"
  )

comp.obj.table.encounters
```

